# Reproducible Research: Peer Assessment 1

## Loading the preprocessing the data ##

We unzip and load the data, downloaded from [the coursera website][data], and
then inspect it for sanity.

```{r, warning=FALSE}
unzip("activity.zip", overwrite=FALSE)
data <- read.csv(file="activity.csv",colClasses=c("integer", "Date", "integer"))
summary(data)
```

We extract two data frames representing the number of steps per day and the mean
steps per time interval.

```{r}
stepsPerDay <- aggregate(steps~date, data=data, sum, na.rm=TRUE)
stepsPerDay <- data.frame(stepsPerDay)

meanStepsPerInterval <- aggregate(steps~interval, data=data, mean, na.rm=TRUE)
meanStepsPerInterval <- data.frame(meanStepsPerInterval)
```

## What is the mean total number of steps taken per day? ##

We use ggplot2 to create a histogram of the average number of steps taken per
day.

```{r, echo=FALSE, eval=FALSE}
## IGNORE THIS CHUNK!!! ##
hist(stepsPerDay,
     breaks=5,
     main="Histograms of steps taken per day",
     xlab="Steps per day",
     ylab="Frequency")
```

```{r, echo=TRUE, eval=TRUE}
library(ggplot2)
qplot(x=steps,
      data=stepsPerDay,
      binwidth=max(stepsPerDay$steps)/5,
      main="Histogram of steps taken per day",
      xlab="Steps per day",
      ylab="Frequency")
```

The **mean** number of steps per day is **10766**, and the **median** number of
steps per day is **10765**.  This was found using the code below, but can also
be found by calling `summary(stepsPerDay)`.

```{r, echo=TRUE, eval=TRUE}
mean(stepsPerDay$steps); median(stepsPerDay$steps)
```

## What is the average daily activity pattern? ##

We make a time series plot of the 5-minute interval and the
average number of steps taken, averaged across all days.

```{r, echo=FALSE, eval=FALSE}
## IGNORE THIS CHUNK!!! ##
plot(interval,
     steps, type="l",
     main="Average number of steps taken per interval",
     xlab="Five-minute Interval",
     ylab="Average steps taken")
```

```{r, echo=TRUE, eval=TRUE}
library(ggplot2)
ggplot(meanStepsPerInterval, aes(x=interval, y=steps)) +
    geom_line() +
    ggtitle("Average number of steps taken per interval") +
    xlab("Interval") +
    ylab("Average steps taken")
```

The five-minute interval starting at **835** represents the most active time,
with ***206.1698*** steps on average.  This was found by calling the code below

```{r, echo=TRUE, eval=TRUE}
meanStepsPerInterval[which.max(meanStepsPerInterval$steps),]
```

## Imputing missing values ##

The total number of missing values in the dataset is **2304**.  This was found
by running the code below, though it an also be found by calling
`summary(data)`.

```{r, echo=TRUE, eval=TRUE}
sum(is.na(data$steps))
```

We replace all NA values in `data$steps` with the mean number of steps per
corresponding interval.

```{r, echo=TRUE, eval=TRUE}
getMeanSteps <- function(interval) {
    meanStepsPerInterval[which(meanStepsPerInterval$interval==interval),]$steps
}
for (i in 1:length(data$steps)) {
    data$steps[i] <- ifelse(is.na(data$steps[i]),
                            getMeanSteps(data$interval[i]),
                            data$steps[i])
}
```

We check that we have successfully removed the NA values in `data$steps`

```{r, echo=TRUE, eval=TRUE}
any(is.na(data$steps))
```

## Are there differences in activity patterns between weekdays and weekends? ##

We create a new factor variable in the dataset with two levels---"weekday"" and
"weekend".

```{r, echo=TRUE, eval=TRUE}
weekday <- weekdays(data$date)
type <- ifelse(weekday == "Saturday" | weekday == "Sunday", "weekend", "weekday")
type <- factor(type)
```

We use ggplot1 to make a panel plot containing a time series plot of the
5-minute interval and the average number of steps taken, averaged
across all weekday days or weekend days.

```{r, echo=FALSE, eval=FALSE}
## IGNORE THIS CHUNK!!! ##
df <- tapply(data$steps, list(data$interval, type), mean)
interval <- as.numeric(row.names(df))

par(mfrow=c(2,1))
plot(interval, df[,"weekday"], type="l",
     main="Weekdays",
     xlab="Interval",
     ylab="Average number of steps")
plot(interval, df[,"weekend"], type="l",
     main="Weekends",
     xlab="Interval",
     ylab="Average number of steps")
```

```{r, echo=TRUE, eval=TRUE}
library(ggplot2)
library(plyr)
df <- ddply(.data=data.frame(data$steps),
            .variables=.(data$interval, type),
            colwise(mean))
names(df) <- c("interval", "type", "stepsPerInterval")
ggplot(df, aes(x=interval, y=stepsPerInterval, xlab="Interval")) +
    geom_line() +
    facet_grid(type~.) +
    ggtitle("Activity pattern on on weekdays versus weekends") +
    xlab("Interval") +
    ylab("Average number of steps")
```

[data]: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip
